Spatial Reasoning and Data Displays

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.Rnw
scaling in biplot

ggbiplot <- function(pca, choices, segColor="black") {
  u <- pca$x[,choices]
  v <- pca$rotation[,choices]
  u <- data.frame(sweep(u, MARGIN=2, 1/pca$sdev[choices], FUN="*"))
  v <- data.frame(sweep(v, MARGIN=2, 1/sqrt(rowSums(v^2))[choices], FUN="*"))
  v$label <- row.names(v)

@@

  qplot(u[,1], u[,2], data=u, colour=I("grey50"), shape=I(1), size=I(3)) + theme_bw() + 
    theme(aspect.ratio=1) + coord_equal() + 
  geom_segment(aes(x=0, y=0, xend=v[,1], yend=v[,2], colour=segColor), data=v, arrow=arrow(length = unit(0.3, "cm"))) + 
    xlab(sprintf("PC%d",choices[1])) + ylab(sprintf("PC%d",choices[2])) +
    geom_text(aes(x=v[,1], y=v[,2], label=label, colour=segColor, 
                  hjust=0.15*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v) +
    scale_colour_identity()
}
@

>>>>

ggbiplot <- function(pca, choices, segColor="black") {
  u <- pca$x[,choices]
  v <- pca$rotation[,choices]
  n <- nrow(u)
  u <- data.frame(sweep(u, MARGIN=2, 1/(pca$sdev[choices]), FUN="*"))
  v <- data.frame(sweep(v, MARGIN=2, 3 * (pca$sdev[choices]), FUN="*"))
  v$label <- row.names(v)

@@

  qplot(u[,1], u[,2], data=u, colour=I("grey50"), shape=I(1), size=I(3)) + theme_bw() + 
    theme(aspect.ratio=1) + coord_equal() + 
  geom_segment(aes(x=0, y=0, xend=v[,1], yend=v[,2], colour=segColor), data=v, arrow=arrow(length = unit(0.3, "cm"))) + 
    xlab(sprintf("PC%d",choices[1])) + ylab(sprintf("PC%d",choices[2])) +
    geom_text(aes(x=v[,1], y=v[,2], label=label, colour=segColor, 
                  hjust=0.15*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v) +
    scale_colour_identity() + theme(aspect.ratio=1)
}
@
